---
title: "Trajectory analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# setup environment
## library
```{r}
library(igraph)
library(mgcv)
library(quadprog) 
library(pcaMethods) 
library(Rcpp) 
library(inline) 
library(RcppArmadillo) 
library(pbapply)
library(glmnet)
library(conos)

library(crestree)

```

## objects

```{r}
conE <- readRDS('/d0/home/jennydongwx/DRG_cochlea/res/Apr2021/conE.rds')
```

```{r}
# extract embedding mat
emb <- conE$embedding

# cluster
conE_clusters <- conE$clusters$leiden$groups

## get the raw counts matrix
raw.mat <- conE$getJointCountMatrix(raw= T)
```

```{r}
# split conchela and DRG cells
cochela.cells <- rownames(emb)[grep(pattern = 'cochlea', x = rownames(emb))]
DRG.cells <- rownames(emb)[grep(pattern = 'DRG', x = rownames(emb))]
```


# building trajectory tree
## DRG
```{r}
# 0.1, 16000
M <- round(dim(emb[DRG.cells, ])[1] / 10)
length(M) # 2161

sigma <- 0.1
lambda <- 160000
metrics <- "euclidean"
z <- ppt.tree(X=t(emb[DRG.cells, ]),M=M, emb=emb, lambda=lambda, sigma=sigma, metrics=metrics, err.cut = 5e-3, n.steps=200, seed=1, plot=F)

# plot embedding
plotppt(z,emb,tips=T,cex.tree = 0.1,cex.main=0.2,lwd.tree = 1)
```

remove spurious branch
```{r}
tmp <- cleanup.branches(z,tips.remove = 1592)
plotppt(tmp,emb,tips=TRUE,forks=FALSE,cex.tree = 0.2,lwd.tree = 2)
ppt <- tmp
```

```{r}
# set root on cells
ppt <- setroot(ppt,root=801)
ppt <- project.cells.onto.ppt(ppt,emb,n.mapping = 5)
```

```{r}
saveRDS(ppt, file = '/d0/home/jennydongwx/DRG_cochlea/res/Apr2021/z_DRG_0.1_16_raw.rds')
```

## cochlea
### filter cochlea cells
cells from non-continous region on the embedding cannot be modeled as part of a progression, thus removed
```{r}
cochlea_clusters <- conE_clusters[cochela.cells]
cochlea_selected <- names(cochlea_clusters[!cochlea_clusters %in% c(8,11)])
length(cochlea_selected) #5886
```

### build tree
```{r}
# final param: 0.05, 40000
M <- round(dim(emb[cochlea_selected, ])[1] / 10)
length(M) # 589

sigma <- 0.05
lambda <- 40000
metrics <- "euclidean"
ppt <- ppt.tree(X=t(emb[cochlea_selected, ]),M=M, emb=emb, lambda=lambda, sigma=sigma, metrics=metrics, err.cut = 5e-3, n.steps=250, seed=1, plot=F)

# plot embedding
plotppt(ppt,emb,tips=T,cex.tree = 0.1,cex.main=0.2,lwd.tree = 1)
```

```{r}
# set root on cells
ppt <- setroot(ppt,root=293)
ppt <- project.cells.onto.ppt(ppt,emb,n.mapping = 5)
```

```{r}
saveRDS(ppt, file = '/d0/home/jennydongwx/DRG_cochlea/res/Apr2021/z_cochlea_0.05_4_raw.rds')
```

# branch DE genes
## DRG
branch DE analysis
```{r}
# look at bifrcation event MSC vs. NMSC
root <- 1736
leaves <- c(270, 332)
subtree <- extract.subtree(ppt,c(root,leaves))
plotppt(ppt,emb,tips=TRUE,forks=FALSE,cex.tree = 0.3,lwd.tree = 3,subtree=subtree)
```

```{r}
fork.de <- test.fork.genes(ppt,t(raw.mat),root=root,leaves=leaves,n.mapping = 1, n.cores = 12)
fork.de <- branch.specific.genes(fork.de,effect.b1 = 0.1,effect.b2 = 0.1, stf.cut = 0)
```

```{r}
# tree DE genes
genes.tree <- ppt$stat.association %>% filter(sign == T)
```

```{r}
DRG_NMSC <- fork.de[intersect(rownames(fork.de)[fork.de$state==2], rownames(genes.tree)), ]
DRG_NMSC$tissue <- 'DRG'
DRG_NMSC$gene <- rownames(DRG_NMSC)

DRG_MSC <- fork.de[intersect(rownames(fork.de)[fork.de$state==1], rownames(genes.tree)), ]
DRG_MSC$tissue <- 'DRG'
DRG_MSC$gene <- rownames(DRG_MSC)
```

```{r}
fork.de['Ncmap', ]$stf >= 0 & fork.de['Ncmap', ]$effect  > 0.1 & fork.de['Ncmap', ]$pd1.a > 0 & fork.de['Ncmap', ]$pd1.p < 0.05
```

## cochlea
branch DE analysis
```{r}
# look at bifrcation event MSC vs. NMSC
root <- 163
leaves <- c(59, 69)
subtree <- extract.subtree(ppt_cochlea,c(root,leaves))
plotppt(ppt_cochlea,emb,tips=TRUE,forks=FALSE,cex.tree = 0.3,lwd.tree = 3,subtree=subtree)
```

```{r}
ppt <- ppt_cochlea
fork.de.cochlea <- test.fork.genes(ppt,t(raw.mat),root=root,leaves=leaves,n.mapping = 1, n.cores = 12)
```

```{r}
fork.de.cochlea <- branch.specific.genes(fork.de.cochlea,effect.b1 = 0.1,effect.b2 = 0.1, stf = 0)
```

```{r}
cochlea_NMSC <- fork.de.cochlea[intersect(rownames(fork.de.cochlea)[fork.de.cochlea$state==2], rownames(genes.tree)), ]
#branch2_up <- fork.de.cochlea %>% filter(stf >= 0 & effect  < -0.1)
#cochlea_NMSC <- fork.de.cochlea[intersect(rownames(branch2_up), rownames(genes.tree)), ]
cochlea_NMSC[1, ] <- rep('placeholder', 10)
cochlea_NMSC$celltype <- 'NMSC'
cochlea_NMSC$gene <- rownames(cochlea_NMSC)

cochlea_MSC <- fork.de.cochlea[intersect(rownames(fork.de.cochlea)[fork.de.cochlea$state==1], rownames(genes.tree)), ]
cochlea_MSC <- fork.de.cochlea[intersect(rownames(branch1_up), rownames(genes.tree)), ]
cochlea_MSC$tissue<- 'cochlea'
cochlea_MSC$gene <- rownames(cochlea_MSC)
```

## take the intersection of DE genes

### Table S1
```{r}
schwann_MSC_idx <- intersect(rownames(DRG_MSC), rownames(cochlea_MSC))
# remove Rik cells
schwann_MSC_idx <- schwann_MSC_idx[!grepl(schwann_MSC_idx, pattern = 'Rik', fixed = TRUE)]
schwann_MSC_idx <- schwann_MSC_idx[!grepl(schwann_MSC_idx, pattern = 'mt-', fixed = TRUE)]

schwann_MSC <- merge(x = DRG_MSC[schwann_MSC_idx,-c(10,11)], y = cochlea_MSC[rownames(cochlea_MSC) %in%schwann_MSC_idx, -c(11,10)], by = 'gene', suffixes = c('.DRG', '.cochlea'))
schwann_MSC$celltype <- 'MSC'
rownames(schwann_MSC) <- schwann_MSC$gene
write.csv(schwann_MSC, file = '/d0/home/jennydongwx/DRG_cochlea/res/Apr2021/MSC_DEG.csv', quote = FALSE)
```

```{r}
schwann_NMSC_idx <- union(rownames(DRG_NMSC), rownames(cochlea_NMSC))
# remove Rik cells
schwann_NMSC_idx <- schwann_NMSC_idx[!grepl(schwann_NMSC_idx, pattern = 'Rik', fixed = TRUE)]
schwann_NMSC_idx <- schwann_NMSC_idx[!grepl(schwann_NMSC_idx, pattern = 'mt-', fixed = TRUE)]

schwann_NMSC <- merge(x = DRG_NMSC[,-c(10,11)], y = cochlea_NMSC[, -c(11,10)], by = 'gene', suffixes = c('.DRG', '.cochlea'))

schwann_NMSC <- DRG_NMSC
schwann_NMSC$celltype <- 'NMSC'

write.csv(schwann_NMSC, file = '/d0/home/jennydongwx/DRG_cochlea/res/Apr2021/NMSC_DEG.csv', quote = FALSE)
```

# plotting

## trajectory tree Fig. 4F

```{r}
plot_segments <- function(tree, color, line_width = 2) {
  al <- get.edgelist(graph.adjacency(tree$B>0))
  al <- matrix(as.integer(al),ncol=2) # adj matrix for branches
  F <- t(t(t(emb[rownames(tree$R),])%*%tree$R)/colSums(tree$R))
  segments(F[1,al[,1]],F[2,al[,1]],F[1,al[,2]],F[2,al[,2]],lwd=line_width, col = color)
}
```

```{r}
r <- z_DRG
vi = rownames(emb)%in%rownames(r$R); names(vi) <- rownames(emb) # logical factor
F <- t(t(t(emb[rownames(r$R),])%*%r$R)/colSums(r$R)) # principal point coord
cols <- rep("grey",nrow(emb)); names(cols) <- rownames(emb) # colors for embedding
pattern.tree <- NULL
tree.col = "black"

# plot
tiff(filename = '/d0/home/jennydongwx/DRG_cochlea/res/Apr2021/final_figures/slide22s.tiff',
     width = 300, height = 300, units = 'mm', res = 200)
     

par(mar=rep(1,4), pty="s")
# background
plot(emb,pch=ifelse(vi,19,1),cex=0.5,col = adjustcolor(cols,ifelse(is.null(pattern.tree),1,0.1)),
     xlab=NA,ylab=NA,xaxt='n',yaxt='n',main=NULL,cex.main=1,font.main=1)


# DRG tree:
plot_segments(z_DRG,  color = 'blue', line_width = 5)

# cochlea tree:
plot_segments(z_cochlea,  color = 'red', line_width = 5)

dev.off()
```


## Fig 4G. DRG (upper)

```{r}
ppt <- readRDS(file = '/d0/home/jennydongwx/DRG_cochlea/res/Apr2021/z_DRG_0.1_16_raw.rds')
```

change color to match cochlea tree
```{r}
color_df <- ppt$cell.summary
color_df <- color_df %>% select(-color) %>% mutate(color = case_when(
  seg == 1 ~ '#FF0000',
  seg == 2 ~ '#00FF00',
  seg == 3 ~ '#0000FF',
  seg == 4 ~ '#CCFF00',
  seg == 5 ~ '#CC00FF',
))
ppt$cell.summary <- color_df

pp_info <- ppt$pp.info
pp_info <- pp_info %>% select(-color) %>% mutate(color = case_when(
  seg == 1 ~ '#FF0000',
  seg == 2 ~ '#00FF00',
  seg == 3 ~ '#0000FF',
  seg == 4 ~ '#CCFF00',
  seg == 5 ~ '#CC00FF',
))
ppt$pp.info <- pp_info
```

### DRG trajectory tree (Fig. 4G upper left)
```{r}
tiff(filename = '/d0/home/jennydongwx/DRG_cochlea/res/Apr2021/final_figures/slide30_L_color.tiff',width = 300, height = 300, units = 'mm', res = 200)
par(mar=rep(1,4), pty="s")
cols <- adjustcolor(ppt$cell.summary[rownames(ppt$R),]$color,0.2); names(cols) <- rownames(ppt$R)
plotppt(ppt,emb,cex.col=cols,
        cex.main=0.5, tree.col=ppt$pp.info$color,cex.title=1,cex.tree=1,lwd.tree=1)
dev.off()
```

### Ncmap (Fig. 4G upper right)
```{r}
tiff(filename = '/d0/home/jennydongwx/DRG_cochlea/res/Apr2021/final_figures/slide30_R.tiff',
     width = 300, height = 300, units = 'mm', res = 200)
par(pty="s")
gene <- 'Ncmap'
plotppt(ppt,emb,pattern.cell = ppt$fit.summary[gene,],gene=gene,cex.main=1,cex.tree = 1.0,lwd.tree = 0.1,par=FALSE, cex.title=6)
dev.off()
```

### Scn7a (Fig. S10 upper right)
```{r}
tiff(filename = '/d0/home/jennydongwx/DRG_cochlea/res/Apr2021/final_figures/slide32_R.tiff',
     width = 300, height = 300, units = 'mm', res = 200)
gene <- 'Scn7a'
par(pty="s")
plotppt(ppt,emb,pattern.cell = ppt$fit.summary[gene,],gene=gene,cex.main=1,cex.tree = 1.0,lwd.tree = 0.1,par=FALSE, cex.title=6)
dev.off()
```

### pseudotime plot Ncmap (Fig. 4G upper middle)

```{r}
 tiff(filename = '/d0/home/jennydongwx/DRG_cochlea/res/Apr2021/final_figures/slide30_M_color.tiff',width = 250, height = 150, units = 'mm', res = 200)
custom_ind <- subset(ppt$cell.summary, !(seg == 1 & t > 3.87))
ppt_removed <- ppt
ppt_removed$cell.summary <- custom_ind
gene <- 'Ncmap'
visualise.trajectory(ppt_removed,gene,t(mat)[gene,],cex.main = 4,lwd.t2=0.5, cex.lab=3)
dev.off()
```

### pseudotime plot Scn7a (Fig. S10 upper middle)

```{r}
# tiff(filename = '/d0/home/jennydongwx/DRG_cochlea/res/Apr2021/final_figures/slide32_M.tiff',
#     width = 250, height = 150, units = 'mm', res = 200)
tiff(filename = '/d0/home/jennydongwx/DRG_cochlea/res/Apr2021/final_figures/slide32_M_color.tiff',
    width = 250, height = 150, units = 'mm', res = 200)
gene <- 'Scn7a'
custom_ind <- subset(ppt$cell.summary, !(v0 == 332 & v1 == 1361))
ppt_removed <- ppt
ppt_removed$cell.summary <- custom_ind
visualise.trajectory(ppt_removed,gene,t(mat)[gene,],cex.main = 4,lwd.t2=0.5, cex.lab=3)
dev.off()
```

## Fig. 4G cochlea (lower)
load cochlea object
```{r}
ppt <- readRDS('/d0/home/jennydongwx/DRG_cochlea/res/Apr2021/z_cochlea_0.05_4_raw.rds')
```


### cochlea trajectory tree (Fig. 4G lower left)
```{r}
tiff(filename = '/d0/home/jennydongwx/DRG_cochlea/res/Apr2021/final_figures/slide3133_L.tiff',
     width = 300, height = 300, units = 'mm', res = 200)
par(mar=rep(1,4), pty="s")
ppt <- project.cells.onto.ppt(ppt,emb,n.mapping = 5)
dev.off()
```

### Ncmap (Fig. 4G lower right)
```{r}
tiff(filename = '/d0/home/jennydongwx/DRG_cochlea/res/Apr2021/final_figures/slide31R.tiff',
     width = 300, height = 300, units = 'mm', res = 200)
gene <- 'Ncmap'
par(pty="s")
plotppt(ppt,emb,pattern.cell = ppt$fit.summary[gene,],gene=gene,cex.main=1,cex.tree = 1.0,lwd.tree = 0.1,par=FALSE, cex.title=6)
dev.off()
```


### Scn7a (Fig. S10 lower right)
```{r}
tiff(filename = '/d0/home/jennydongwx/DRG_cochlea/res/Apr2021/final_figures/slide33R.tiff',
     width = 300, height = 300, units = 'mm', res = 200)
gene <- 'Scn7a'
par(pty="s")
plotppt(ppt,emb,pattern.cell = ppt$fit.summary[gene,],gene=gene,cex.main=1,cex.tree = 1.0,lwd.tree = 0.1,par=FALSE, cex.title=6)
dev.off()
```

### pseudotime plot Ncmap (Fig. 4G lower middle)

```{r}
tiff(filename = '/d0/home/jennydongwx/DRG_cochlea/res/Apr2021/final_figures/slide31_M.tiff',
     width = 250, height = 150, units = 'mm', res = 200)
gene <- 'Ncmap'
custom_ind <- subset(ppt$cell.summary, !(v0 == 59 & v1 == 98))
ppt_removed <- ppt
ppt_removed$cell.summary <- custom_ind
visualise.trajectory(ppt_removed,gene,t(mat)[gene,],cex.main = 4,lwd.t2=0.5, cex.lab=3)
dev.off()
```

### pseudotime plot Scn7a (Fig. S10 lower middle)
```{r}
tiff(filename = '/d0/home/jennydongwx/DRG_cochlea/res/Apr2021/final_figures/slide33_M.tiff',
     width = 250, height = 150, units = 'mm', res = 200)
gene <- 'Scn7a'
custom_ind <- subset(ppt$cell.summary, !(v0 == 69 & v1 == 449))
ppt_removed <- ppt
ppt_removed$cell.summary <- custom_ind
visualise.trajectory(ppt_removed,gene,t(mat)[gene,],cex.main = 4,lwd.t2=0.5, cex.lab=3)
dev.off()
```
